---
title: "Metanalysis Solution"
author: "Husnain Mustafa"
date: today
format:
  html:
    code-fold: true
    code-summary: "Show the code"
execute:
  echo: true
  warning: false
theme: darkly
---

## Loading Data and Packages:
Start by loading the data and verifying that all the requirements are met.
```{r}
# Install and load necessary packages

library(readxl)
library(meta)
library(metafor)
library(ggplot2)

# Import the data from the Excel file
data <- read_excel("../data/metaanalysis_data.xlsx")

# View the first few rows of the data
head(data)
```

Next we can generate some aditional variables to help us perform teh required analysis. 

## Effect Sizes for Toys:
Calculating individual effect sizes for Male and Female toys allongwith their variane. 
```{r}
# Calculate effect sizes and variances for male toys
data$effect_size_male <- data$Mean_boys_play_male - data$Mean_girls_play_male
data$variance_male <- (data$SD_boys_play_male^2 / data$N_boys) + (data$SD_girls_play_male^2 / data$N_girls)

# Calculate effect sizes and variances for female toys
data$effect_size_female <- data$Mean_boys_play_female - data$Mean_girls_play_female
data$variance_female <- (data$SD_boys_play_female^2 / data$N_boys) + (data$SD_girls_play_female^2 / data$N_girls)

```
## Perfroming Meta Analysis:
Now we can perform the meta analysis on the effect sizes for male and female toys.

### Male Toys
```{r}
# Meta-analysis for male toys
res_male <- rma(yi = data$effect_size_male, vi = data$variance_male, method="REML")

# Summary of the meta-analysis for male toys
summary(res_male)

```

```{r}
# Funnel plot for male toys
funnel(res_male)

# Optional: Enhanced funnel plot with ggplot2
funnel_plot_male <- ggplot(data, aes(x = effect_size_male, y = sqrt(variance_male))) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Funnel Plot for Male Toys", x = "Effect Size", y = "Standard Error") +
  theme_minimal()

print(funnel_plot_male)

```

### Female Toys
```{r}
# Meta-analysis for female toys
res_female <- rma(yi = data$effect_size_female, vi = data$variance_female, method="REML")

# Summary of the meta-analysis for female toys
summary(res_female)
```

```{r}
# Funnel plot for female toys
funnel(res_female)

# Optional: Enhanced funnel plot with ggplot2
funnel_plot_female <- ggplot(data, aes(x = effect_size_female, y = sqrt(variance_female))) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Funnel Plot for Female Toys", x = "Effect Size", y = "Standard Error") +
  theme_minimal()

print(funnel_plot_female)

```
### Adding Aditional Variables
Meta-regression for male toys, assuming quality columns are binary indicators (0 or 1).
Summary of the meta-regression for male toys
```{r}
res_meta_reg_male <- rma(yi = data$effect_size_male, vi = data$variance_male, 
                         mods = ~ data$`Neutral toys` + data$`Parent present`  + data$Setting + data$Country + data$`NOS score`,
                         method="REML")


summary(res_meta_reg_male)

```

Meta-regression for female toys.
Summary of the meta-regression for female toys.
```{r}

res_meta_reg_female <- rma(yi = data$effect_size_female, vi = data$variance_female, 
                           mods = ~ data$`Neutral toys` + data$`Parent present`  + data$Setting + data$Country + data$`NOS score`,
                           method="REML")


summary(res_meta_reg_female)

```

Meta-regression to check if author gender affects the results for male toys
```{r}

res_author_gender_male <- rma(yi = data$effect_size_male, vi = data$variance_male, 
                              mods = ~ data$`Female authors` + data$`Male authors`,
                              method="REML")


summary(res_author_gender_male)

```

Meta-regression to check if author gender affects the results for female toys
```{r}

res_author_gender_female <- rma(yi = data$effect_size_female, vi = data$variance_female, 
                                mods = ~ data$`Female authors` + data$`Male authors`,
                                method="REML")

# Summary of the meta-regression for female toys
summary(res_author_gender_female)

```

